---
title: 'Differential Abundance methods- Simulation Design'
author: 'Saritha Kodikara'
#date: 'November 9, 2017'
date: '\today'
#date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  pdf_document:
    toc: true    # can remove
    toc_depth: 3 # can remove
    number_sections: true
  highlight: zenburn
header-includes:
- \usepackage{fancyhdr}
- \usepackage{mathtools}
- \usepackage{xcolor, hyperref}
- \usepackage{lipsum}
- \usepackage{caption}
- \usepackage{diagbox}
- \usepackage{multirow}
- \setlength{\headheight}{28pt}
- \setlength{\footskip}{25pt}
- \pagestyle{fancy}
- \renewcommand{\headrulewidth}{0.5pt}
- \renewcommand{\footrulewidth}{0.5pt}
- \lhead{\includegraphics[width=8cm,height=1cm]{logo-unimelb}}
- \cfoot{\scriptsize MAST30033 Statistical Genomics | School of Mathematics and Statistics \\ The University of Melbourne}
- \rhead{\thepage}
- \hypersetup{colorlinks   = true, linkcolor=blue, urlcolor  = blue}
- \fancypagestyle{plain}{\pagestyle{fancy}}
editor_options: 
  chunk_output_type: console
bibliography: packages.bib
---

```{r global_options, include=FALSE}

.libPaths("/data/projects/punim0613/Saritha/R/lib_4.0.0")
lib= .libPaths()[1]
library(knitr)
# global options to show by default the code, dump the figures into /Figures etc
knitr::opts_chunk$set(dpi = 100, 
                      echo=TRUE, 
                      warning=FALSE, message=FALSE, eval = TRUE,
                      fig.show=TRUE,fig.align='center', fig.path= 'Figures/')
library(ggplot2)
library(dplyr)
library(tscount)
library(patchwork)
```


<!-- Start your report from here. -->


<!-- the {-} indicate that you do not wish to number this section -->
# Simulation Design

To assess the performance of the differential abundance methods discussed in this review paper, we simulated longitudinal count data from a generalized linear model with a negative binomial distribution using the 'tscount' R package [@liboschik2017tscount]. We estimated realistic parameter values for dispersion and Autoregressive (AR) coefficient to be used in the simulation using the pregnancy data [@DiGiulio2015]. Based on these estimated values nine case scenarios were created with three dispersion (i.e., noise) values [0.1;0.3;0.6] and three AR values [0.04; 0.2; 0.4]. 


\begin{table*}[t]
  \centering
  \caption{Parameter values used in scenario 1 for 300 taxa.}
    \begin{tabular}{lcccc}
          \multicolumn{1}{p{16.75em}}{\textbf{Effect - Number of taxa}} & \multicolumn{1}{l}{\textbf{Intercept}} & \multicolumn{1}{l}{\textbf{Time}} & \multicolumn{1}{l}{\textbf{Group}} & \multicolumn{1}{l}{\textbf{Group*Time}} \\
          Time - 10 & Uniform(0,5) & 1.5   & 0     & 0 \\
          Group - 10 & Uniform(0,5) & 0     & 13     & 0 \\
          Group+Time+Group*Time -10 & Uniform(0,5) & 1.5   & 13     & 5 \\
          No - 270  & Uniform(0,5) & 0     & 0     & 0 \\
    \end{tabular}%
  \label{tab:simPara}%
\end{table*}%


```{r}
#Simulation function
simulate<-function(mod,para, nIndiv, nTime,disper, nTaxa, nSc1, nSc2, nSc3, meta){
  
  TAXA<-setNames(data.frame(matrix(ncol = nTaxa, 
                nrow = nIndiv*nTime)),paste0("Taxa_",1:nTaxa))
  
  for(i in 1:nTaxa){
    if(i<=nSc1){#Time
      k=1
    }else if(i<=nSc1+nSc2){#Group
      k=2
    }else if(i<=nSc1+nSc2+nSc3){#Time+Group+Time*Group
      k=3
    }else {#No
      k=4
    }
    count<-c()
    for(j in 1:nIndiv){
      t<-c(tsglm.sim(n=nTime, param = para[[k]], model=model, 
                     xreg=matrix(c(1:nTime,
                     rep(meta$Group[meta$Indiv==j][1],nTime),
                     1:nTime*rep(meta$Group[meta$Indiv==j][1],nTime)),ncol=3), 
                     link="identity",
                     distr="nbinom", distrcoefs=c(size=1/disper))$ts)
      count<-c(count,t)
    }
    while(!any(count==0)){#At least one zero (Needed to run ZIGMM)
      count<-c()
      for(j in 1:nIndiv){
        t<-c(tsglm.sim(n=nTime, param = para[[k]], model=model, 
                       xreg=matrix(c(1:nTime,rep(meta$Group[meta$Indiv==j][1],nTime),
                            1:nTime*rep(meta$Group[meta$Indiv==j][1],nTime)),ncol=3),
                       link="identity",
                       distr="nbinom", distrcoefs=c(size=1/disper))$ts)
        count<-c(count,t)
      } 
    }
    TAXA[,i]<-count
  }
  
  
  return(TAXA)
}

```


## Data Simulation- Scenario 1 

In this section, we simulate data under Scenario 1 (i.e., dispersion=0.1; AR value=0.04). For Scenario 1, 50 data sets were simulated with 300 profiles (i.e. taxa) measured on 10-time points and twenty individuals belonging to two groups. These 300 profiles consisted of five profiles that were differentially expressed trough time only, five profiles that were differentially expressed trough group only, five profiles that were differentially expressed trough time and group and 270 profiles that were without any differentially expressed effect (i.e., noise profiles) (see Table \ref{tab:simPara}). All these profiles had at-least one zero count between time 0 and  time 10 for at-least one individual. This condition was included into the simulation as to avoid errors in ZIGMM and FZINBMM model fitting. 

```{r}
set.seed(1234)
model <- list(past_obs=1) #Only 1 AR parameter
disp_1<-0.1
param_1 <- list(list(intercept=runif(1,0,5), past_obs=0.04,  xreg=c(1.5,0,0)),
                list(intercept=runif(1,0,5), past_obs=0.04,  xreg=c(0,13,0)),
                list(intercept=runif(1,0,5), past_obs=0.04,  xreg=c(1.5,13,5)),
                list(intercept=runif(1,0,5), past_obs=0.04,  xreg=c(0,0,0)))

nIndiv=20;nTime=10
metaDF<-data.frame(Time=rep(c(1:nTime),nIndiv),
                   Indiv=rep(1:nIndiv,each=nTime),
                   Group=rep(0:1,each=nIndiv*nTime/2))

n_iter <- 50 # Number of iterations of the loop
list_of_frames <- replicate(n_iter, data.frame())
# Initializes the progress bar
pb <- txtProgressBar(min = 0,      # Minimum value of the progress bar
                     max = n_iter, # Maximum value of the progress bar
                     style = 3,    # Progress bar style
                     width = 50,   # Progress bar width
                     char = "=")   # Character used to create the bar

for(i in 1:n_iter) {
  
  #---------------------
  # Code to be executed
  #---------------------
  df_Scenario1<-simulate(model,param_1,nIndiv,nTime, disp_1, 300, 10,10,10,metaDF)
  list_of_frames[[i]] <-df_Scenario1
  #---------------------
  
  # Sets the progress bar to the current state
  setTxtProgressBar(pb, i)
}

close(pb) # Close the connection

c_Sc1<-lapply(list_of_frames, function(x) {
  mutate(x,Library_size=rowSums(x))
})

ra_Sc1<-lapply(c_Sc1, function(x) {
  x[,-301]/x$Library_size
})

saveRDS(metaDF,file = "Data/df_meta.Rdata")
saveRDS(c_Sc1,file = "Data/count_Scenario1.Rdata")
saveRDS(ra_Sc1,file = "Data/RA_Scenario1.Rdata")
```


```{r sc1plots, fig.show='hold', fig.cap="Figure 1, Some simulated taxa profiles with time (A,a); group (B,b); group * time (C,c); and no effect (D,d) with 0.1 dispersion and 0.04 AR. Top panel shows the counts, while the bottom shows the relative abundance values", fig.fullwidth=TRUE}

dfcount<- cbind(metaDF,c_Sc1[[1]])
dfRA<- cbind(metaDF,ra_Sc1[[1]])


p1<-dfcount %>%
  ggplot(aes(x=Time, y=Taxa_1,  colour= as.factor(Group),
             group =  as.factor(Indiv), shape= as.factor(Group), 
             linetype = as.factor(Group)))+
  geom_line()+ geom_point()+ggtitle("a")+ylab("Count")+
  labs( linetype = "Group", shape = "Group", color="Group") +
  scale_color_manual(values=c("steelblue","darkviolet"))+
  theme(legend.position = "none")

p11<-dfcount %>%
  ggplot(aes(x=Time, y=Taxa_11,  colour= as.factor(Group),
             group =  as.factor(Indiv), shape= as.factor(Group), 
             linetype = as.factor(Group)))+
  geom_line()+ geom_point()+ggtitle("b")+ylab("Count")+
  labs( linetype = "Group", shape = "Group", color="Group")  +
  scale_color_manual(values=c("steelblue","darkviolet"))+
  theme(legend.position = "none")

p21<-dfcount %>%
  ggplot(aes(x=Time, y=Taxa_21,  colour= as.factor(Group),
             group =  as.factor(Indiv), shape= as.factor(Group), 
             linetype = as.factor(Group)))+
  geom_line()+ geom_point()+ggtitle("c")+ylab("Count")+
  labs( linetype = "Group", shape = "Group", color="Group")  +
  scale_color_manual(values=c("steelblue","darkviolet"))+
  theme(legend.position = "none")

p31<-dfcount %>%
  ggplot(aes(x=Time, y=Taxa_31,  colour= as.factor(Group),
             group =  as.factor(Indiv), shape= as.factor(Group), 
             linetype = as.factor(Group)))+
  geom_line()+ geom_point()+ggtitle("d")+ylab("Count")+
  labs( linetype = "Group", shape = "Group", color="Group")  +
  scale_color_manual(values=c("steelblue","darkviolet"))+
  theme(legend.position = "none")


plot1<-p1|p11|p21|p31

p1.1<-dfRA %>%
  ggplot(aes(x=Time, y=Taxa_1,   colour= as.factor(Group),
             group =  as.factor(Indiv), shape= as.factor(Group), 
             linetype = as.factor(Group)))+
  geom_line()+ geom_point()+ggtitle("A")+ylab("Relative Abundance")+
  labs( linetype = "Group", shape = "Group", color="Group")  +
  scale_color_manual(values=c("steelblue","darkviolet"))+
  theme(legend.position = "none")


p11.1<-dfRA %>%
  ggplot(aes(x=Time, y=Taxa_11,  colour= as.factor(Group),
             group =  as.factor(Indiv), shape= as.factor(Group),
             linetype = as.factor(Group)))+
  geom_line()+ geom_point()+ggtitle("B")+ylab("Relative Abundance")+
  labs( linetype = "Group", shape = "Group", color="Group")  +
  scale_color_manual(values=c("steelblue","darkviolet"))+
  theme(legend.position = "none")

p21.1<-dfRA %>%
  ggplot(aes(x=Time, y=Taxa_21,   colour= as.factor(Group),
             group =  as.factor(Indiv), shape= as.factor(Group), 
             linetype = as.factor(Group)))+
  geom_line()+ geom_point()+ggtitle("C")+ylab("Relative Abundance")+
  labs( linetype = "Group", shape = "Group", color="Group")  +
  scale_color_manual(values=c("steelblue","darkviolet"))+
  theme(legend.position = "none")

p31.1<-dfRA %>%
  ggplot(aes(x=Time, y=Taxa_31,   colour= as.factor(Group),
        group =  as.factor(Indiv), shape= as.factor(Group),
        linetype = as.factor(Group)))+
  geom_line()+ geom_point()+ggtitle("D")+ylab("Relative Abundance")+
  labs( linetype = "Group", shape = "Group", color="Group") +
  scale_color_manual(values=c("steelblue","darkviolet"))+
  theme(legend.position = "none")


(p1|p11|p21|p31)/(p1.1|p11.1|p21.1|p31.1)
```

<!-- New page -->
\pagebreak


# References {-}

\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\noindent
<div id="refs"></div>
```{r refmgr references, results="asis", echo=FALSE}
# Print
```
\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}
\setlength{\parskip}{0pt}
