---
title: 'Clustering methods- Results'
author: 'Saritha Kodikara'
#date: 'November 9, 2017'
date: '\today'
#date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  pdf_document:
    toc: true    # can remove
    toc_depth: 3 # can remove
    number_sections: true
  highlight: zenburn
header-includes:
- \usepackage{fancyhdr}
- \usepackage{mathtools}
- \usepackage{xcolor, hyperref}
- \usepackage{lipsum}
- \usepackage{caption}
- \usepackage{diagbox}
- \usepackage{multirow}
- \setlength{\headheight}{28pt}
- \setlength{\footskip}{25pt}
- \pagestyle{fancy}
- \renewcommand{\headrulewidth}{0.5pt}
- \renewcommand{\footrulewidth}{0.5pt}
- \lhead{\includegraphics[width=8cm,height=1cm]{logo-unimelb}}
- \cfoot{\scriptsize School of Mathematics and Statistics \\ The University of Melbourne}
- \rhead{\thepage}
- \hypersetup{colorlinks   = true, linkcolor=blue, urlcolor  = blue}
- \fancypagestyle{plain}{\pagestyle{fancy}}
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}

.libPaths("/data/projects/punim0613/Saritha/R/lib")
lib= .libPaths()[1]
library(knitr)
# global options to show by default the code, dump the figures into /Figures etc
knitr::opts_chunk$set(dpi = 100, 
                      echo=TRUE, 
                      warning=FALSE, message=FALSE, eval = TRUE,
                      fig.show=TRUE,fig.align='center', fig.path= 'Figures/')
source("Source/Utility_functions.R")
library(timeOmics)
library(tidyverse)
library(lmms)
library(magrittr)
library(reshape2)
library(cluster)
library(dtwclust)
library(patchwork)
```


<!-- Start your report from here. -->


<!-- the {-} indicate that you do not wish to number this section -->
# Simulation Results- Original data

In here, we present the clustering accuracy for four clustering method ("PCA", "DTW", "Agglomerative", "Kmedoids") with original linear mixed model (LMM) profiles.

```{r}
temp <-list.files(path="Data/", pattern="*.Rdata")

for(i in 1:3){
  clu_output<-data.frame()
  modelledData<-readRDS(paste0("Data/",temp[i]))
  
  for(j in 1:length(modelledData)){
    lmmData<- modelledData[[j]]
    methods_accuracy<-cluster_Methods_acc(lmmData)
    clu_output<-rbind(clu_output,methods_accuracy)
    names(clu_output)<-c("PCA", "DTW", 
                       "Agglomerative", "Kmedoids")
  }
  write.csv(clu_output, file =  paste0("Results/Org_",i,".csv"))
}


```

```{r clus_Or, fig.show='hold', fig.cap="Figure 1, Clustering accuracy for PCA, K-medoid, DTW and Agglomerative clustering for original LMMS data.", fig.fullwidth=TRUE, fig.width=10, fig.height=8}
temp <-list.files(path="Results/",pattern= "Org_")
for(i in 1:3){
  clRes<-read.csv(paste0("Results/",temp[i]))[,-1]
  nam <- paste("plot", i, sep = "_")
  assign(nam, clusBoxplot(clRes, plotTitle = ""))
}

P<-plot_1+plot_2+plot_3+ plot_layout(guides = "collect")


temp <-list.files(path="Data/", pattern="*.Rdata")
for(i in 1:3){
  clData<-readRDS(paste0("Data/",temp[i]))[[2]]
  nam <- paste("plot", i, sep = "_")
  if(i==1) n="Noise=0.5"
  if(i==2) n="Noise=1.5"
  if(i==3) n="Noise=3"
  assign(nam, plotFunction(clData, plotTitle = n))
}

Q<-plot_1+plot_2+plot_3+ plot_layout(guides = "collect")

Q/P
```


# Simulation Results- Centered data

In here, we present the clustering accuracy for four clustering method ("PCA", "DTW", "Agglomerative", "Kmedoids") with centered LMM profiles.

```{r}
temp <-list.files(path="Data/", pattern="*.Rdata")

for(i in 1:3){
  clu_output<-data.frame()
  modelledData<-readRDS(paste0("Data/",temp[i]))
  
  for(j in 1:length(modelledData)){
    Centered.lmmData<-scale(modelledData[[j]],scale=FALSE)
    methods_accuracy<-cluster_Methods_acc(Centered.lmmData)
    clu_output<-rbind(clu_output,methods_accuracy)
    names(clu_output)<-c("PCA", "DTW", 
                       "Agglomerative", "Kmedoids")
  }
  write.csv(clu_output, file =  paste0("Results/ Center_",i,".csv"))
}


```

```{r clus_Cen, fig.show='hold', fig.cap="Figure 2, Clustering accuracy for PCA, K-medoid, DTW and Agglomerative clustering for centered LMMS data.", fig.fullwidth=TRUE, fig.width=10, fig.height=8}
temp <-list.files(path="Results/",pattern="Center_")
for(i in 1:3){
  clRes<-read.csv(paste0("Results/",temp[i]))[,-1]
  nam <- paste("plot", i, sep = "_")
  assign(nam, clusBoxplot(clRes, plotTitle = ""))
}

P<-plot_1+plot_2+plot_3+ plot_layout(guides = "collect")


temp <-list.files(path="Data/", pattern="*.Rdata")
for(i in 1:3){
  clData<-readRDS(paste0("Data/",temp[i]))[[2]]
  nam <- paste("plot", i, sep = "_")
  if(i==1) n="Noise=0.5"
  if(i==2) n="Noise=1.5"
  if(i==3) n="Noise=3"
  assign(nam, plotFunction(clData, plotTitle = n,center = TRUE))
}

Q<-plot_1+plot_2+plot_3+ plot_layout(guides = "collect")

Q/P
```


# Simulation Results- Scaled data

In here, we present the clustering accuracy for four clustering method ("PCA", "DTW", "Agglomerative", "Kmedoids") with scaled LMM profiles.

```{r}
temp <-list.files(path="Data/", pattern="*.Rdata")

for(i in 1:3){
  clu_output<-data.frame()
  modelledData<-readRDS(paste0("Data/",temp[i]))
  
  for(j in 1:length(modelledData)){
    Scaled.lmmData<-scale(modelledData[[j]],center =FALSE)
    methods_accuracy<-cluster_Methods_acc(Scaled.lmmData)
    clu_output<-rbind(clu_output,methods_accuracy)
    names(clu_output)<-c("PCA", "DTW", 
                       "Agglomerative", "Kmedoids")
  }
  write.csv(clu_output, file =  paste0("Results/ Scale_",i,".csv"))
}


```

```{r clus_Scal, fig.show='hold', fig.cap="Figure 2, Clustering accuracy for PCA, K-medoid, DTW and Agglomerative clustering for scaled LMMS data.", fig.fullwidth=TRUE, fig.width=10, fig.height=8}
temp <-list.files(path="Results/",pattern="Scale_")
for(i in 1:3){
  clRes<-read.csv(paste0("Results/",temp[i]))[,-1]
  nam <- paste("plot", i, sep = "_")
  assign(nam, clusBoxplot(clRes, plotTitle = ""))
}

P<-plot_1+plot_2+plot_3+ plot_layout(guides = "collect")


temp <-list.files(path="Data/", pattern="*.Rdata")
for(i in 1:3){
  clData<-readRDS(paste0("Data/",temp[i]))[[2]]
  nam <- paste("plot", i, sep = "_")
  if(i==1) n="Noise=0.5"
  if(i==2) n="Noise=1.5"
  if(i==3) n="Noise=3"
  assign(nam, plotFunction(clData, plotTitle = n, scale = TRUE))
}

Q<-plot_1+plot_2+plot_3+ plot_layout(guides = "collect")

Q/P
```


# Simulation Results- Centered and Scaled data

In here, we present the clustering accuracy for four clustering method ("PCA", "DTW", "Agglomerative", "Kmedoids") with centered and scaled LMM profiles.

```{r}
temp <-list.files(path="Data/", pattern="*.Rdata")

for(i in 1:3){
  clu_output<-data.frame()
  modelledData<-readRDS(paste0("Data/",temp[i]))
  
  for(j in 1:length(modelledData)){
    CenteredScaled.lmmData<-scale(modelledData[[j]])
    methods_accuracy<-cluster_Methods_acc(CenteredScaled.lmmData)
    clu_output<-rbind(clu_output,methods_accuracy)
    names(clu_output)<-c("PCA", "DTW", 
                       "Agglomerative", "Kmedoids")
  }
  write.csv(clu_output, file =  paste0("Results/ Cen_Scal_",i,".csv"))
}


```

```{r clus_Cen_Scal, fig.show='hold', fig.cap="Figure 2, Clustering accuracy for PCA, K-medoid, DTW and Agglomerative clustering for centered and scaled LMMS data.", fig.fullwidth=TRUE, fig.width=10, fig.height=8}
temp <-list.files(path="Results/",pattern="Cen_Scal_")
for(i in 1:3){
  clRes<-read.csv(paste0("Results/",temp[i]))[,-1]
  nam <- paste("plot", i, sep = "_")
  assign(nam, clusBoxplot(clRes, plotTitle = ""))
}

P<-plot_1+plot_2+plot_3+ plot_layout(guides = "collect")


temp <-list.files(path="Data/", pattern="*.Rdata")
for(i in 1:3){
  clData<-readRDS(paste0("Data/",temp[i]))[[2]]
  nam <- paste("plot", i, sep = "_")
  if(i==1) n="Noise=0.5"
  if(i==2) n="Noise=1.5"
  if(i==3) n="Noise=3"
  assign(nam, plotFunction(clData, plotTitle = n, scale = TRUE, center = TRUE))
}

Q<-plot_1+plot_2+plot_3+ plot_layout(guides = "collect")

Q/P
```