---
title: 'Differential Abundance methods- Simulation Results'
author: 'Saritha Kodikara'
#date: 'November 9, 2017'
date: '\today'
#date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  pdf_document:
    toc: true    # can remove
    toc_depth: 3 # can remove
    number_sections: true
  highlight: zenburn
header-includes:
- \usepackage{fancyhdr}
- \usepackage{mathtools}
- \usepackage{xcolor, hyperref}
- \usepackage{lipsum}
- \usepackage{caption}
- \usepackage{diagbox}
- \usepackage{multirow}
- \setlength{\headheight}{28pt}
- \setlength{\footskip}{25pt}
- \pagestyle{fancy}
- \renewcommand{\headrulewidth}{0.5pt}
- \renewcommand{\footrulewidth}{0.5pt}
- \lhead{\includegraphics[width=8cm,height=1cm]{logo-unimelb}}
- \cfoot{School of Mathematics and Statistics \\ The University of Melbourne}
- \rhead{\thepage}
- \hypersetup{colorlinks   = true, linkcolor=blue, urlcolor  = blue}
- \fancypagestyle{plain}{\pagestyle{fancy}}
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(dpi = 100, 
                      echo=TRUE, 
                      warning=FALSE, message=FALSE, eval = TRUE,
                      fig.show=TRUE, fig.path= 'Figures/')
library(ggplot2)
library(magrittr)
library(dplyr)
library(tscount)
library(ZIBR)
library("NBZIMM")
library(splinectomeR)
library(epiR)
library(stringr)
library(tidyr)
library(patchwork)
library(grid)
library(gridExtra)
library(nlme)
library(BiocParallel)
```


<!-- Start your report from here. -->
# Simulation Results- Scenario 1 

In this file, we present results for Taxa 1 simulated under moderate parameter values (i.e. dispersion=0.1 and AR=0.04) from Zero-inflated Beta Regression Model, Negative Binomial Mixed Model, splinectomeR, Zero-inflated Gaussian mixed model and Fast zero-inflated negative binomial mixed model.

```{r}
meta_df<-readRDS("Data/df_meta.Rdata")
list_sc1_ra<-readRDS("Data/RA_Scenario1.Rdata")
list_sc1_co<-readRDS("Data/count_Scenario1.Rdata")
dfcount<- cbind(meta_df,list_sc1_co[[1]])
dfRA<- cbind(meta_df,list_sc1_ra[[1]])
```


## Zero-inflated Gaussian mixed model

```{r, results='hide'}
#Scenario 1 (Relative Abundance)
fit_zigmm<-lme.zig(Taxa_1~Time+Group+Time*Group, random = ~ 1 | Indiv, data = dfRA)
summary(fit_zigmm)

#Scenario 1 (Relative Abundance)-with AR
fit_zigmmAR<-lme.zig(Taxa_1~Time+Group+Time*Group, random = ~ 1 | Indiv, 
                     data = dfRA, correlation = corAR1())
summary(fit_zigmmAR)

#Scenario 1 (Counts)
fit_zigmm_count<-lme.zig(Taxa_1~Time+Group+Time*Group+ offset(log(Library_size)),
                    random = ~ 1 | Indiv, data = dfcount)
summary(fit_zigmm_count)

#Scenario 1 (Counts)-with AR
fit_zigmmAR_count<-lme.zig(Taxa_1~Time+Group+Time*Group+ offset(log(Library_size)),
                    random = ~ 1 | Indiv, data = dfcount, correlation = corAR1())
summary(fit_zigmmAR_count)
```

## Zero-inflated Beta Regression Model

```{r, results='hide'}
#Scenario 1
df_cov = data.frame(Time = meta_df$Time, Group=meta_df$Group, 
                      Time_Group=meta_df$Time*meta_df$Group)
fit_zibr<-zibr(logistic.cov = data.frame(Time = meta_df$Time, Group=meta_df$Group, 
                      Time_Group=meta_df$Time*meta_df$Group), 
                     beta.cov = df_cov, 
                     Y =list_sc1_ra[[1]]$Taxa_1, subject.ind = meta_df$Indiv,
                     time.ind =  meta_df$Time)
fit_zibr
```

## Negative Binomial Mixed Model

```{r, results='hide'}
#Scenario 1
fit_nbmm<-glmm.nb(Taxa_1~Time+Group+Time*Group+ offset(log(Library_size)),
                    random = ~ 1 | Indiv, data = dfcount)
summary(fit_nbmm)

#Scenario 1-AR
fit_nbmmAR<-glmm.nb(Taxa_1~Time+Group+Time*Group+ offset(log(Library_size)),
                    random = ~ 1 | Indiv, data = dfcount, correlation = corAR1())
summary(fit_nbmmAR)
```

## Fast zero-inflated negative binomial mixed model

```{r, results='hide'}
#Scenario 1
fit_zinb<-glmm.zinb(Taxa_1~Time+Group+Time*Group+ offset(log(Library_size)),
                    random = ~ 1 | Indiv, data = dfcount)
summary(fit_zinb)

#Scenario 1-AR
fit_zinbAR<-glmm.zinb(Taxa_1~Time+Group+Time*Group+ offset(log(Library_size)),
                    random = ~ 1 | Indiv, data = dfcount, correlation = corAR1())
summary(fit_zinbAR)

```

## SplinectomeR

```{r, results='hide'}

#Scenario 1
# Test for difference in RA change over time between groups 0 and 1
fit_permuspliner <- permuspliner(data =  dfRA, x = 'Time', y = 'Taxa_1',
                       cases = 'Indiv', category = 'Group', groups = c('0','1'))
fit_permuspliner$pval

# Test for non-zero trend in RA over Time
fit_trendyspliner <- trendyspliner(data = dfRA, x = 'Time', y = 'Taxa_1',
                        cases = 'Indiv', perms = 999)
fit_trendyspliner$pval
```

